---
title: "Model Building"
output: html_notebook
---


```{r}
library(dplyr)       # Data manipulation
library(ggplot2)     # Data visualization
library(readr)       # CSV file reading
#install.packages("corrplot")
library(corrplot)    # Correlation plot
library(stringr)
library(leaps) # subset selection package
```

# Data exploration and variable transformation (dummies/label encoder)
```{r}
# Load the dataset
df <- read.csv("hdb_cleaned_2022onwards.csv")
```

```{r}
print(df)
```
```{r}
colnames(df)
```
```{r}
# drop all unnecessary columns that will not be relevant for price prediction
df = subset(df, select = -c(X, month, block, street_name, lease_commence_date, address, latitude, longitude,
                            school_shortest_dist_name, good_sch_list_within_1km,
                            mrt_shortest_dist_name, mrt_list_within_1km, mall_shortest_dist_name, 
                            mall_list_within_1km, communitycentre_shortest_dist_name, 
                            communitycentre_list_within_1km, eldercare_shortest_dist_name, eldercare_list_within_1km,
                            hawker_shortest_dist_name, hawker_list_within_1km, supermarket_shortest_dist_name,
                            supermarket_list_within_1km))
df
```
```{r}
colnames(df)
```
Storey Range
```{r}
table(df$storey_range)[order(names(table(df$storey_range)))]
```
```{r}
ggplot(df, aes(x = storey_range)) +
  geom_histogram(binwidth = 1,stat = "count", fill = "blue", color = "black") +
  labs(title = "Histogram of Storey Range", x = "Storey Range", y = "Frequency")
```
```{r}
## label encoding for Storey Range
# The higher the storey_range, the higher the label encoding

# Perform label encoding

# Sort the data frame by 'storey_range' before label encoding
df = df[order(df$storey_range), ]

# Perform label encoding using the factor function
df$storey_range_label <- as.numeric(factor(df$storey_range))

# Display the resulting data frame
print(df[df$storey_range == "49 TO 51",])
```
Flat Type
```{r}
ggplot(df, aes(x = flat_type)) +
  geom_histogram(binwidth = 1,stat = "count", fill = "blue", color = "black") +
  labs(title = "Histogram of Flat Type", x = "Flat Type", y = "Frequency")
```
```{r}
table(df$flat_type)[order(names(table(df$flat_type)))]
```

```{r}
# Sort the data frame by 'flat_type' before label encoding
df = df[order(df$flat_type), ]

# Perform label encoding using the factor function
df$flat_type_label <- as.numeric(factor(df$flat_type))

# Display the resulting data frame
print(df[df$flat_type == "EXECUTIVE",])
```
Flat Model
```{r}
ggplot(df, aes(x = flat_model)) +
  geom_histogram(binwidth = 1,stat = "count", fill = "blue", color = "black") +
  labs(title = "Histogram of Flat Model", x = "Flat Model", y = "Frequency") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}
table(df$flat_model)
```
```{r}
# Re-categorize flat models into broader categories
replace_values = c("Model A-Maisonette" = "Maisonette", "Improved-Maisonette" = "Maisonette", "Premium Maisonette" = "Maisonette")
df$flat_model = str_replace_all(df$flat_model, replace_values)

table(df$flat_model)
```

```{r}
# get dummies 
dummies = model.matrix(~ flat_model-1, data = df)

df = cbind(df, dummies)

```

```{r}
dummies
```

Town
```{r}
ggplot(df, aes(x = town)) +
  geom_histogram(binwidth = 1,stat = "count", fill = "blue", color = "black") +
  labs(title = "Histogram of Town", x = "Town", y = "Frequency") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}
# get dummies
# get dummies 
dummies = model.matrix(~ town - 1, data = df)

df = cbind(df, dummies)
df
```
Create subset df to contain only columns for machine learning training purposes
```{r}
df_ml = df = subset(df, select = -c(town, flat_type, storey_range, flat_model,good_sch_score_sap_gep,good_sch_score))
df_ml
```
```{r}
## Do we need to ln to change distribution to normal?
ggplot(df_ml, aes(x = log(resale_price))) +
  geom_histogram(bin = 100, stat = 'count', fill = "blue", color = "black") +
  labs(title = "Histogram of Ln Resale Price", x = "Ln Resale Price", y = "Frequency") 
```
```{r}
#add ln(resale_price)
df_ml['ln_resale_price'] = log(df_ml['resale_price'])
```


```{r}
colnames(df_ml)
```
```{r fig.height=9, fig.width=12}
correlation_matrix <- cor(df_ml %>% select(resale_price, floor_area_sqm, remaining_lease, school_shortest_dist, good_sch_count_within_1km,
                                        mrt_shortest_dist, mall_shortest_dist,mrt_count_within_1km,
                                          mall_shortest_dist, mall_count_within_1km,communitycentre_shortest_dist,communitycentre_count_within_1km,
                                          eldercare_shortest_dist,eldercare_count_within_1km,hawker_shortest_dist,hawker_count_within_1km,
                                          supermarket_shortest_dist,supermarket_count_within_1km,distance_to_cbd_km,matured, ln_resale_price))
corrplot(correlation_matrix, 
  method = "color",     # Color method for the plot
   addCoef.col="black", 
   order = "AOE", 
   number.cex=0.75)

```
df_ml should contain all the relevant columns fit for ML training. Might still need to do feature scaling/normalization of predictors, or maybe need to do log transformation to any variables if we need to adhere to normality assumptiobs
```{r}
colnames(df_ml)
```


```{r}
write.csv(df_ml, 'df_ml.csv', row.names=FALSE)
```


